Testing gene methylation enrichment analysis approaches using simulated data

### 2023-10-15 figs/GSE158422_simulate.Rmd
simwork

@@

# divide sets between hyper and hypomethylated
nset <- floor(length(gset_mod)/2)
set.seed(seed) ; gtup <-sample(gset_mod,48)
set.seed(seed) ; gtdn <- sample(setdiff(gset_mod,gtup),nset)
gup <- unname(unlist(gtup))
gdn <- unname(unlist(gtdn))

@@

# select probes

>>>>

frac_genes=0.5
frac_probes=0.5

@@

# divide gene sets between hyper and hypomethylated
nset <- floor(length(gset_mod)/2)
set.seed(seed) ; gtup <-sample(gset_mod,48)
set.seed(seed) ; gtdn <- sample(setdiff(gset_mod,gtup),nset)
gup <- unname(unlist(gtup))
gdn <- unname(unlist(gtdn))

# make probe-gene vector
probe2gene <- strsplit(myann$UCSC_RefGene_Name,";")
names(probe2gene) <- rownames(myann)
probe2gene <- unlist(probe2gene)

# select probes hypermethylated
set.seed(seed) ; gup2 <- sample(gup,floor(length(gup)/2))
pup <- names(probe2gene[which(probe2gene %in% gup2)])
set.seed(seed) ; pup2 <- sample(pup,floor(length(pup)/2))

# select probes hypomethylated
set.seed(seed) ; gdn2 <- sample(gdn,floor(length(gdn)/2))
pdn <- names(probe2gene[which(probe2gene %in% gdn2)])
set.seed(seed) ; pdn2 <- sample(pdn,floor(length(pup)/2))

@@

# incorporate altered signals - change by +1 or -1
mval_case[rownames(mval_case) %in% pup2,] <-  mval_case[rownames(mval_case) %in% pup2,] +1
mval_case[rownames(mval_case) %in% pdn2,] <-  mval_case[rownames(mval_case) %in% pdn2,] -1
mval2 <- cbind(mval_ctrl,mval_case)

ss <- data.frame(colnames(mval2))
colnames(ss) <- "sample"
ss$case <- c(rep(0,18),rep(1,18))

d <- model.matrix(~ ss$case)
fit.reduced <- lmFit(mval2,d)
fit.reduced <- eBayes(fit.reduced)

summary(decideTests(fit.reduced))
dm <- topTable(fit.reduced,coef=ncol(d), number = Inf)
